---
title: "Labor and Delivery Dashboard"
format: 
  dashboard:
    nav-buttons: [github]
    github: https://github.com/mine-cetinkaya-rundel/ld-dashboard
logo: images/logo.png
theme: [sandstone, theme/custom.scss]
fig-width: 10
fig-asp: 0.3
params:
  month: "October"
  year: "2023"
  # 2021 rates: https://www.cdc.gov/nchs/data/nvsr/nvsr72/nvsr72-01.pdf
  us_cesarean_rate: 0.321 
  us_preterm_rate:  0.1049
  threshold_diff: 0.02
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(readxl)
library(scales)
library(DT)
library(gt)
theme_set(theme_minimal(base_size = 24, base_family = "Atkinson Hyperlegible"))
```

```{r}
#| label: load-data
#| message: false

ld <- read_excel("data/ld.xlsx")
```

```{r}
#| label: set-inputs

time_period <- paste(params$month, params$year)
```

```{r}
#| label: prep-data

ld <- ld |>
  mutate(
    delivery_method = str_to_sentence(delivery_method),
    term = str_to_sentence(term),
    parity = case_when(
      parity == "1"  ~ "Nulliparous",
      parity == "2"  ~ "Primiparous",
      parity == "3+" ~ "Multiparous"
    ),
    parity = fct_relevel(parity, "Nulliparous", "Primiparous", "Multiparous"),
    maternal_age = fct_relevel(maternal_age, "<20", "20-34", "35+")
  )
```

#  {.sidebar}

This dashboard displays statistics for:

|              |                     |
|--------------|---------------------|
| **Hospital** | Grey Sloan Memorial |
| **Unit**     | Labor and Delivery  |
| **Month**    | `{r} time_period`   |

------------------------------------------------------------------------

In `{r} time_period` the staff breakdown in the unit was as follows:

|                          |     |
|--------------------------|-----|
| **Attending physicians** |  14 |
| **Residents**            |  21 |
| **Nurses**               |  12 |

------------------------------------------------------------------------

::: {.callout-note collapse="true"}
## Disclaimer

This is a fictional hospital. The data are simulated based on realistic birth characteristics and risk factors from [this report by the CDC](https://www.cdc.gov/nchs/data/nvsr/nvsr72/nvsr72-01.pdf).
:::

# All

```{r}
#| label: all-values
#| results: hide

n_births <- nrow(ld)

p_cesarean <- ld |>
  count(delivery_method) |>
  mutate(p = n / sum(n)) |>
  filter(delivery_method == "Cesarean") |>
  pull(p)

p_cesarean_color <- case_when(
  between(p_cesarean, params$us_cesarean_rate, params$us_cesarean_rate + params$threshold_diff) ~ "warning",
  p_cesarean > params$us_cesarean_rate + params$threshold_diff ~ "danger",
  .default = "light"
  )

p_preterm <- ld |>
  count(term) |>
  mutate(p = n / sum(n)) |>
  filter(term == "Pre-term") |>
  pull(p)

p_preterm_color <- case_when(
  between(p_preterm, params$us_preterm_rate, params$us_preterm_rate + params$threshold_diff) ~ "warning",
  p_preterm > params$us_preterm_rate + params$threshold_diff ~ "danger",
  .default = "light"
  )
```

## Row {height="20%"}

```{r}
#| content: valuebox
#| title: "Total births"

list(
  icon = "file-medical",
  color = "primary",
  value = n_births
)
```

```{r}
#| content: valuebox
#| title: "Cesarean deliveries"

list(
  icon = "clipboard2-pulse",
  color = p_cesarean_color,
  value = label_percent(accuracy = 0.1)(p_cesarean)
)
```

```{r}
#| content: valuebox
#| title: "Pre-term births"

list(
  icon = "calendar-week",
  color = p_preterm_color,
  value = label_percent(accuracy = 0.1)(p_preterm)
)
```

## Row {height="40%"}

### Column {width="40%"}

```{r}
#| title: Delivery method

ld |>
  count(delivery_method) |>
  mutate(p = n / sum(n)) |>
  gt() |>
  fmt_percent(
    columns = p,
    decimals = 1
  ) |>
  tab_style(
    style = cell_text(color = "#ae8b2d", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = delivery_method == "Cesarean"
    )
  ) |>
  tab_style(
    style = cell_text(color = "#0e2635", weight = "bold"),
    locations = cells_body(
      columns = everything(),
      rows = delivery_method == "Vaginal"
    )
  ) |>
  cols_label(
    delivery_method = "",
    n = "Number of<br>deliveries",
    p = "Proportion of<br>deliveries",
    .fn = md
  )
```

### Column {width="60%"}

```{r}
#| title: Maternal age

ld |>
  count(maternal_age) |>
  mutate(p = n / sum(n)) |>
  ggplot(aes(x = maternal_age, y = n, group = 1)) +
  geom_line(linewidth = 2, color = "#0e263560") +
  geom_point(size = 3, color = "#0e2635") +
  geom_point(size = 6, color = "#0e2635", shape = "circle open") +
  labs(x = NULL) +
  scale_y_continuous(
    "Count",
    sec.axis = sec_axis(~ . / n_births, name = "Proportion", labels = label_percent())
  )
```

## Row {height="40%"}

```{r}
#| title: Delivery method and parity

ld |>
  count(parity, delivery_method) |>
  ggplot(aes(x = n, y = fct_rev(parity), fill = delivery_method)) +
  geom_col(position = "fill", color = "white") +
  scale_fill_manual(
    values = c("#ae8b2d", "#0e2635"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(labels = label_percent()) +
  labs(y = NULL, x = NULL, fill = "Delivery\nmethod")
```

```{r}
#| title: Delivery method and maternal age

ld |>
  count(maternal_age, delivery_method) |>
  ggplot(aes(x = n, y = fct_rev(maternal_age), fill = delivery_method)) +
  geom_col(position = "fill", color = "white") +
  scale_fill_manual(
    values = c("#ae8b2d", "#0e2635"),
    guide = guide_legend(reverse = TRUE)
  ) +
  scale_x_continuous(labels = label_percent()) +
  labs(y = NULL, x = NULL, fill = "Delivery\nmethod")
```

# Vaginal {orientation="columns"}

## Column {width="60%"}

```{r}
#| label: vaginal-values
#| results: hide

ld_v <- ld |>
  filter(delivery_method == "Vaginal")

n_births_v <- nrow(ld_v)

p_preterm_v <- ld_v |>
  count(term) |>
  mutate(p = n / sum(n)) |>
  filter(term == "Pre-term") |>
  pull(p)

p_preterm_color_v <- case_when(
  between(p_preterm_v, params$us_preterm_rate, params$us_preterm_rate + params$threshold_diff) ~ "warning",
  p_preterm_v > params$us_preterm_rate + params$threshold_diff ~ "danger",
  .default = "light"
  )
```

### Row {height="20%"}

```{r}
#| component: valuebox
#| title: "Total births"

list(
  icon = "file-medical",
  color = "primary",
  value = n_births_v
)
```

```{r}
#| component: valuebox
#| title: "Pre-term births"

list(
  icon = "calendar-week",
  color = p_preterm_color_v,
  value = label_percent(accuracy = 0.1)(p_preterm_v)
)
```

### Row {height="40%"}

```{r}
#| title: Maternal age

ld_v |>
  count(maternal_age) |>
  mutate(p = n / sum(n)) |>
  ggplot(aes(x = maternal_age, y = n, group = 1)) +
  geom_line(linewidth = 2, color = "#0e263560") +
  geom_point(size = 3, color = "#0e2635") +
  geom_point(size = 6, color = "#0e2635", shape = "circle open") +
  labs(x = NULL) +
  scale_y_continuous(
    "Count",
    sec.axis = sec_axis(~ . / n_births, name = "Proportion", labels = label_percent())
  )
```

### Row {height="40%" .tabset}

```{r}
#| title: Maternal age and parity

ld_v |>
  count(maternal_age, parity) |>
  ggplot(aes(x = n, y = fct_rev(maternal_age), fill = fct_rev(parity))) +
  geom_col(position = "fill", color = "white") +
  scale_fill_brewer(
    palette = "Blues",
    guide = guide_legend(reverse = TRUE),
    direction = -1
  ) +
  scale_x_continuous(labels = label_percent()) +
  labs(y = NULL, x = NULL, fill = "Parity")
```

```{r}
#| title: Maternal age and term

ld_v |>
  count(maternal_age, term) |>
  ggplot(aes(x = n, y = fct_rev(maternal_age), fill = fct_rev(term))) +
  geom_col(position = "fill", color = "white") +
  scale_fill_brewer(
    palette = "Greens",
    guide = guide_legend(reverse = TRUE),
    direction = -1
  ) +
  scale_x_continuous(labels = label_percent()) +
  labs(y = NULL, x = NULL, fill = "Term")
```

## Column {width="40%"}

```{r}
#| title: Data

ld_v |>
  select(-delivery_method) |>
  arrange(patient_id) |>
  datatable(
    colnames = c("ID", "Maternal age", "Parity", "Term"),
    options = list(
      dom = "lrtip", 
      paging = TRUE,
      lengthMenu = c(5, 10, 25),
      pageLength = 10, 
      scrollY = TRUE
    ),
  )
```

# Caesarean {orientation="columns"}

## Column {width="60%"}

```{r}
#| label: cesarean-values
#| results: hide

ld_c <- ld |>
  filter(delivery_method == "Cesarean")

n_births_c <- nrow(ld_c)

p_preterm_c <- ld_c |>
  count(term) |>
  mutate(p = n / sum(n)) |>
  filter(term == "Pre-term") |>
  pull(p)

p_preterm_color_c <- case_when(
  between(p_preterm_c, params$us_preterm_rate, params$us_preterm_rate + params$threshold_diff) ~ "warning",
  p_preterm_c > params$us_preterm_rate + params$threshold_diff ~ "danger",
  .default = "light"
  )
```

### Row {height="20%"}

```{r}
#| component: valuebox
#| title: "Total births"

list(
  icon = "file-medical",
  color = "primary",
  value = n_births_c
)
```

```{r}
#| component: valuebox
#| title: "Pre-term births"

list(
  icon = "calendar-week",
  color = p_preterm_color_c,
  value = label_percent(accuracy = 0.1)(p_preterm_c)
)
```

### Row {height="40%"}

```{r}
#| title: Maternal age

ld_c |>
  count(maternal_age) |>
  mutate(p = n / sum(n)) |>
  ggplot(aes(x = maternal_age, y = n, group = 1)) +
  geom_line(linewidth = 2, color = "#0e263560") +
  geom_point(size = 3, color = "#0e2635") +
  geom_point(size = 6, color = "#0e2635", shape = "circle open") +
  labs(x = NULL) +
  scale_y_continuous(
    "Count",
    sec.axis = sec_axis(~ . / n_births, name = "Proportion", labels = label_percent())
  )
```

### Row {height="40%" .tabset}

```{r}
#| title: Maternal age and parity

ld_c |>
  count(maternal_age, parity) |>
  ggplot(aes(x = n, y = fct_rev(maternal_age), fill = fct_rev(parity))) +
  geom_col(position = "fill", color = "white") +
  scale_fill_brewer(
    palette = "Blues",
    guide = guide_legend(reverse = TRUE),
    direction = -1
  ) +
  scale_x_continuous(labels = label_percent()) +
  labs(y = NULL, x = NULL, fill = "Parity")
```

```{r}
#| title: Maternal age and term

ld_c |>
  count(maternal_age, term) |>
  ggplot(aes(x = n, y = fct_rev(maternal_age), fill = fct_rev(term))) +
  geom_col(position = "fill", color = "white") +
  scale_fill_brewer(
    palette = "Greens",
    guide = guide_legend(reverse = TRUE),
    direction = -1
  ) +
  scale_x_continuous(labels = label_percent()) +
  labs(y = NULL, x = NULL, fill = "Term")
```

## Column {width="40%"}

```{r}
#| title: Data

ld_c |>
  select(-delivery_method) |>
  arrange(patient_id) |>
  datatable(
    colnames = c("ID", "Maternal age", "Parity", "Term"),
    options = list(
      dom = "lrtip", 
      paging = TRUE,
      lengthMenu = c(5, 10, 25),
      pageLength = 10, 
      scrollY = TRUE
    ),
  )
```

# Data

```{r}
ld |>
  arrange(patient_id) |>
  datatable(
    colnames = c("ID", "Maternal age", "Delivery method", "Parity", "Term"),
    options = list(dom = 'ftp', paging = TRUE)
    )
```
